---
title: 'Homework 1 Answers'
fig-cap: false
editor: 
  markdown: 
    wrap: 72
---

-   If you find any typos or errors, please feel free to contact me via
    email at r13323002\@ntu.edu.tw. I would appreciate it and will
    correct the solution to prevent any misunderstandings.

-   If you?e having trouble understanding the code, don? hesitate to
    ask me for help. I?l do my best to assist you during office hours
    or after the TA session.


## Computer Exercises

[(census.csv)](census.csv){target="_blank" rel="noopener"}

``` r
# clear the memory and import package
rm(list = ls())
library(tidyverse)

# import data
census <- read.csv("census.csv")
```

```{r}
#| echo: false

# clear the memory and import package
rm(list = ls())
library(tidyverse)

# import data
census <- read.csv("census.csv")
```

### Summary Statistics Table

``` r
summary_table <- data.frame(
  mean = c(
    mean(census$total_personal_income, na.rm = TRUE),
    mean(census$total_family_income, na.rm = TRUE),
    mean(census$age, na.rm = TRUE)
  ),
  sd = c(
    sd(census$total_personal_income, na.rm = TRUE),
    sd(census$total_family_income, na.rm = TRUE),
    sd(census$age, na.rm = TRUE)
  ),
  min = c(
    min(census$total_personal_income, na.rm = TRUE),
    min(census$total_family_income, na.rm = TRUE),
    min(census$age, na.rm = TRUE)
  ),
  q25 = c(
    quantile(census$total_personal_income, 0.25, na.rm = TRUE),
    quantile(census$total_family_income, 0.25, na.rm = TRUE),
    quantile(census$age, 0.25, na.rm = TRUE)
  ),
  median = c(
    median(census$total_personal_income, na.rm = TRUE),
    median(census$total_family_income, na.rm = TRUE),
    median(census$age, na.rm = TRUE)
  ),
  q75 = c(
    quantile(census$total_personal_income, 0.75, na.rm = TRUE),
    quantile(census$total_family_income, 0.75, na.rm = TRUE),
    quantile(census$age, 0.75, na.rm = TRUE)
  ),
  max = c(
    max(census$total_personal_income, na.rm = TRUE),
    max(census$total_family_income, na.rm = TRUE),
    max(census$age, na.rm = TRUE)
  ),
  row.names = c("total_personal_income", "total_family_income", "age")
)

round(summary_table, 2)
```

```{r}
#| echo: false

summary_table <- data.frame(
  mean = c(
    mean(census$total_personal_income, na.rm = TRUE),
    mean(census$total_family_income, na.rm = TRUE),
    mean(census$age, na.rm = TRUE)
  ),
  sd = c(
    sd(census$total_personal_income, na.rm = TRUE),
    sd(census$total_family_income, na.rm = TRUE),
    sd(census$age, na.rm = TRUE)
  ),
  min = c(
    min(census$total_personal_income, na.rm = TRUE),
    min(census$total_family_income, na.rm = TRUE),
    min(census$age, na.rm = TRUE)
  ),
  q25 = c(
    quantile(census$total_personal_income, 0.25, na.rm = TRUE),
    quantile(census$total_family_income, 0.25, na.rm = TRUE),
    quantile(census$age, 0.25, na.rm = TRUE)
  ),
  median = c(
    median(census$total_personal_income, na.rm = TRUE),
    median(census$total_family_income, na.rm = TRUE),
    median(census$age, na.rm = TRUE)
  ),
  q75 = c(
    quantile(census$total_personal_income, 0.75, na.rm = TRUE),
    quantile(census$total_family_income, 0.75, na.rm = TRUE),
    quantile(census$age, 0.75, na.rm = TRUE)
  ),
  max = c(
    max(census$total_personal_income, na.rm = TRUE),
    max(census$total_family_income, na.rm = TRUE),
    max(census$age, na.rm = TRUE)
  ),
  row.names = c("total_personal_income", "total_family_income", "age")
)

round(summary_table, 2)
```

### Distribution of Total Personal Income

``` r
census_income <- census[!is.na(census$total_personal_income), ]

ggplot(census_income, aes(x = total_personal_income)) +
  geom_histogram(binwidth = 10000, fill = "steelblue", color = "black") +
  labs(
    x = "Total Personal Income",
    y = "Frequency",
    title = "Distribution of Total Personal Income"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#| fig-width: 8
#| fig-height: 5
#| out-width: "90%"
#| fig-align: center
#| echo: false

census_income <- census[!is.na(census$total_personal_income), ]

ggplot(census_income, aes(x = total_personal_income)) +
  geom_histogram(binwidth = 10000, fill = "steelblue", color = "black") +
  labs(
    x = "Total Personal Income",
    y = "Frequency",
    title = "Distribution of Total Personal Income"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

The distribution of total personal income is strongly right-skewed. Most
observations are concentrated in the lower income ranges, especially from
$0$ to about $40{,}000$. There are also several high-income observations,
with the maximum equal to $456{,}000$, which appear as potential outliers.

### Age and Total Personal Income

``` r
ggplot(census_income, aes(x = age, y = total_personal_income)) +
  geom_point(color = "steelblue") +
  labs(
    x = "Age",
    y = "Total Personal Income",
    title = "Age and Total Personal Income"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#| fig-width: 8
#| fig-height: 5
#| out-width: "90%"
#| fig-align: center
#| echo: false

ggplot(census_income, aes(x = age, y = total_personal_income)) +
  geom_point(color = "steelblue") +
  labs(
    x = "Age",
    y = "Total Personal Income",
    title = "Age and Total Personal Income"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

There appears to be a weak positive relationship between age and personal
income. The points are widely scattered, so the association is not strong.

### Personal Income by Sex

``` r
ggplot(census_income, aes(x = sex, y = total_personal_income)) +
  geom_boxplot(fill = "steelblue") +
  labs(
    x = "Sex",
    y = "Total Personal Income",
    title = "Total Personal Income by Sex"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#| fig-width: 8
#| fig-height: 5
#| out-width: "90%"
#| fig-align: center
#| echo: false

ggplot(census_income, aes(x = sex, y = total_personal_income)) +
  geom_boxplot(fill = "steelblue") +
  labs(
    x = "Sex",
    y = "Total Personal Income",
    title = "Total Personal Income by Sex"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

The male group has a higher median personal income. The male group also
exhibits greater overall variability, with a wider interquartile range and
more extreme high-income observations.

### Sex and Race Contingency Table

``` r
sex_race_table <- table(census$sex, census$race_general)
sex_race_table_withsum <- addmargins(sex_race_table)
sex_race_table_withsum
```

```{r}
#| echo: false

sex_race_table <- table(census$sex, census$race_general)
sex_race_table_withsum <- addmargins(sex_race_table)
sex_race_table_withsum
```

``` r
ggplot(census, aes(x = race_general, fill = sex)) +
  geom_bar(position = "dodge") +
  coord_flip() +
  labs(
    x = "General Race",
    y = "Headcount",
    fill = "Sex",
    title = "Headcount by Sex and Race"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#| fig-width: 8
#| fig-height: 5
#| out-width: "90%"
#| fig-align: center
#| echo: false

ggplot(census, aes(x = race_general, fill = sex)) +
  geom_bar(position = "dodge") +
  coord_flip() +
  labs(
    x = "General Race",
    y = "Headcount",
    fill = "Sex",
    title = "Headcount by Sex and Race"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

The largest race category in this sample is White for both sexes. The bar
plot shows that male and female counts are broadly similar within most race
categories, though males are slightly more numerous in the full sample.
## Textbook Exercises

### Question 1.4

a. The research question is: "Do asthmatic patients who practice the Buteyko method experience improvement in their condition?"

b. The cases are 600 adult patients aged 18--69 years diagnosed and currently treated for asthma.

c. The variables are whether or not the patient practiced the Buteyko method (categorical) and measures of quality of life, activity, asthma symptoms, and medication reduction of the patients (categorical, ordinal). It may also be reasonable to treat the ratings on a scale of 0 to 10 as discrete numerical variables.

### Question 1.10

a. Each row of the data matrix represents a participant in the survey.

b. There are 1,691 participants in the survey.

c. The table below lists the types of variables included in the study:

| Variable      | Type                   |
|:------------- |:---------------------- |
| sex           | categorical            |
| age           | numerical, continuous  |
| maritalStatus | categorical            |
| grossIncome   | ordinal\*              |
| smoke         | categorical            |
| amtWeekends   | numerical, discrete    |
| amtWeekdays   | numerical, discrete    |

\* Note that income is a continuous numerical variable, but in this survey, it is recorded as an ordinal variable.

### Question 1.22

a. Observational.  
b. No. Because the study is observational, we cannot infer causation.  
c. Possible confounders: caffeine intake and lack of sleep. Either could explain the observed association between stress and muscle cramps (e.g., students may cramp due to increased caffeine use or insufficient sleep).

### Question 1.38

a. Simple random sampling: effective -- gives every household equal probability of selection.  
b. Stratified sampling: effective -- neighborhoods are distinct; stratifying ensures sampling from each neighborhood.  
c. Cluster sampling: not effective -- sampling whole neighborhoods would miss households from some areas, and neighborhoods differ greatly.  
d. Multi-stage sampling: suffers from the same issue as (c).  
e. Convenience sampling: not effective -- yields a biased sample (e.g., only nearby/similar neighborhoods; misses distant ones).

### Question 1.40

a. Explanatory variable: percent of the population with a bachelor's degree.  
   Response variable: per-capita income (in thousands).  
b. Relationship: strong positive linear association -- as the bachelor's-degree percent increases, per-capita income tends to increase. Few counties exceed 60% bachelor's degree or $50k per-capita income.  
c. Causality: observational study; no causal claim. We can only say higher bachelor's-degree share is associated with higher per-capita income.

### Question 2.12

Median is around 80. Because the distribution is left-skewed, the mean is expected to be slightly lower than the median.

### Question 2.30

a. If $\dfrac{\bar{x}}{\mathrm{median}} = 1$, then $\bar{x}=\mathrm{median}$. This matches symmetric distributions.  
b. If $\dfrac{\bar{x}}{\mathrm{median}} < 1$, then $\bar{x}<\mathrm{median}$. This matches left-skewed distributions (lower values pull the mean down).  
c. If $\dfrac{\bar{x}}{\mathrm{median}} > 1$, then $\bar{x}>\mathrm{median}$. This matches right-skewed distributions (higher values pull the mean up).