---
title: 'Homework 2 Answers'
fig-cap: false
editor: 
  markdown: 
    wrap: 72
---

-   If you find any typos or errors, please feel free to contact me via
    email at r13323002\@ntu.edu.tw. I would appreciate it and will
    correct the solution to prevent any misunderstandings.

-   For some R commands marked with ‘OPTIONAL,’ you don’t need to learn
    them. You will be completely fine on the computer quiz even if you
    choose to ignore them.

-   If you’re having trouble understanding the code, don’t hesitate to
    ask me for help. I’ll do my best to assist you during office hours
    or after the TA session.

## Textbook Exercises

### Exercise 4.1

a.  

$$
\widehat{\text{TestScore}} = 640.3 - 4.93 \times 25 = 517.05
$$

b.  

$$
\begin{aligned}
\Delta \widehat{\text{TestScore}}
&= (640.3 - 4.93 \times 24) - (640.3 - 4.93 \times 21) \\
&= -4.93 \times (24 - 21) \\
&= -14.79
\end{aligned}
$$

c.  

$$
\begin{aligned}
\overline{\text{TestScore}}
&= \hat{\beta}_0 + \hat{\beta}_1 \times \overline{\text{CS}} \\
&= 640.3 - 4.93 \times 22.8 \\
&= 527.896
\end{aligned}
$$

d.  

$$
\begin{aligned}
s_Y = \sqrt{\frac{1}{n - 1} TSS}
&= \sqrt{\frac{1}{n - 1} \cdot \frac{SSR}{1 - R^2}} \\
&= \sqrt{\frac{1}{n - 1} \cdot \frac{(n-2)SER^2}{1 - R^2}} \\
&= \sqrt{\frac{1}{50 - 1} \cdot \frac{(50 - 2) \cdot (8.7)^2}{1 - 0.11}} \approx 9.1274
\end{aligned}
$$

### Exercise 4.3

a.  

-   **Intercept:** A worker with €0 average monthly income has a
    predicted average monthly expenditure of €710.7.
-   **Slope:** AME is expected to increase by €8.8 for each additional
    euro of income.

b.  SER is in the same units as the dependent variable. That is, SER is
    measured in euros (or dollars) per month.

c.  $R^2$ is unit free.

d.  If $AMI = 100$, then

$$
AME = 710.7 + 8.8 \times 100 = 1590.7
$$ If $AMI = 200$, then

$$
AME = 710.7 + 8.8 \times 200 = 2470.7
$$

e.  No. The highest income in the sample is €1.5 million, and €2 million
    is far outside the range of the sample data.

f.  No. The distribution of earnings is positively skewed.

### Exercise 4.7

$$
\begin{aligned}
E\left(\hat{\beta}_0\right)
&= E\left(\overline{Y} - \hat{\beta}_1 \overline{X}_1\right) \\
&= E \left[ \left( \beta_0 + \beta_1 \overline{X}_1 + \frac{1}{n} \sum_{i=1}^{n} u_i \right) - \hat{\beta}_1 \overline{X}_1 \right] \\
&= E \left[ \beta_0 + (\beta_1 - \hat{\beta}_1) \overline{X}_1 + \frac{1}{n} \sum_{i=1}^{n} u_i \right] \\
&= E (\beta_0) + E \left\{ E \left[ (\beta_1 - \hat{\beta}_1) \overline{X}_1 \mid X_1 \right] \right\} + E \left(\frac{1}{n} \sum_{i=1}^{n} u_i\right) \\
&= \beta_0 + E \left[ \overline{X}_1 \underbrace{E \left(\beta_1 - \hat{\beta}_1 \mid X_1\right)}_{=0} \right] + \frac{1}{n} \sum_{i=1}^{n} \underbrace{E(u_i)}_{=0} \\
&= \beta_0
\end{aligned}
$$

### Exercise 4.12

a.  Note that

$$
\begin{aligned}
    ESS
    &= \sum_{i=1}^{n} (Y_i - \bar{Y})^2 \\
    &= \sum_{i=1}^{n} (\hat{\beta}_0 + \hat{\beta}_1 X_i - \bar{Y})^2 \\
    &= \sum_{i=1}^{n} [\hat{\beta}_1 (X_i - \bar{X})]^2 \\
    &= \hat{\beta}_1^2 \sum_{i=1}^{n} (X_i - \bar{X})^2
    \end{aligned}
$$ Then

$$
\begin{aligned}
    R^2 &= \frac{ESS}{TSS} \\
    &= \frac{\hat{\beta}_1^2 \sum_{i=1}^{n} (X_i - \overline{X})^2}{\sum_{i=1}^{n} (Y_i - \overline{Y})^2} \\
    &= \left[ \frac{\sum_{i=1}^{n} (X_i - \overline{X}) (Y_i - \overline{Y})}{\sum_{i=1}^{n} (X_i - \overline{X})^2} \right]^2
    \cdot \frac{\sum_{i=1}^{n} (X_i - \overline{X})^2}{\sum_{i=1}^{n} (Y_i - \overline{Y})^2} \\
    &= \frac{\left[\sum_{i=1}^{n} (X_i - \overline{X}) (Y_i - \overline{Y})\right]^2}
    {\sum_{i=1}^{n} (X_i - \overline{X})^2 \sum_{i=1}^{n} (Y_i - \overline{Y})^2} \\
    &= \left[ \frac{\sum_{i=1}^{n} (X_i - \overline{X}) (Y_i - \overline{Y})}
    {\sqrt{\sum_{i=1}^{n} (X_i - \overline{X})^2} \sqrt{\sum_{i=1}^{n} (Y_i - \overline{Y})^2}} \right]^2 \\
    &= \left[ \frac{\frac{1}{n-1} \sum_{i=1}^{n} (X_i - \overline{X}) (Y_i - \overline{Y})}
    {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n} (X_i - \overline{X})^2}
    \sqrt{\frac{1}{n-1} \sum_{i=1}^{n} (Y_i - \overline{Y})^2}} \right]^2 \\
    &= \left[ \frac{s_{XY}}{s_X \cdot s_Y} \right]^2 \\
    &= r_{XY}^2
    \end{aligned}
$$

b.  Since $r_{XY} = r_{YX}$, the statement is true.

c.  

$$
\begin{aligned}
r_{XY} \cdot \frac{s_Y}{s_X} &= \frac{s_{XY}}{s_X \cdot s_Y} \cdot \frac{s_Y}{s_X} \\
&= \frac{\frac{1}{n-1} \sum_{i=1}^{n} (X_i - \overline{X}) (Y_i - \overline{Y})}
{\frac{1}{n-1} \sum_{i=1}^{n} (X_i - \overline{X})^2} \\
&= \frac{\sum_{i=1}^{n} (X_i - \overline{X}) (Y_i - \overline{Y})}
{\sum_{i=1}^{n} (X_i - \overline{X})^2} = \hat{\beta}_1.
\end{aligned}
$$ 


### Exercise 5.2

a.  The estimated gender wage gap is **\$1.78** per hour.

b.  **Prepare**:

$$
\left\{\begin{aligned}
    &H_0: \beta_{\text{Male}} = 0 \\
    &H_1: \beta_{\text{Male}} \neq 0
    \end{aligned}\right.
$$ Let the significance level be $\alpha = 0.05$.

**Calculate**:

$$
t^* = \frac{1.78 - 0}{0.29} \approx 6.14
$$ $$
p\text{-value} = 2 \Phi(-t^*) \approx 2 \Phi(-6.14) \approx 0
$$ **Conclude**:

```         
Since $p$-value < 0.05, we reject $H_0$. There is significant evidence that the gender wage gap is different from 0.
```

c.  Confidence Interval:

$$
\hat{\beta}_{\text{Male}} \pm Z_{0.025} SE(\hat{\beta}_{\text{Male}}) = 1.78 \pm 1.96 \times 0.29 = [1.21, 2.35]
$$

d.  The sample average wage of women is

$$
\hat{\beta}_0 = 10.73 \quad \text{(in dollars per hour)}
$$ The sample average wage of men is

$$
\hat{\beta}_0 + \hat{\beta}_{\text{Male}} = 10.73 + 1.78 = 12.51 \quad \text{(in dollars per hour)}
$$

e.  Regression equation:

$$
\begin{aligned}
    \text{Wage}
    &= (\hat{\beta}_0 + \hat{\beta}_{\text{Male}}) + (-\hat{\beta}_{\text{Male}}) \times \text{Female} \\
    &= 12.51 - 1.78 \times \text{Female}, \quad R^2 = 0.09, \quad SER = 3.8
    \end{aligned}
$$

### Exercise 5.5

a.  The estimated gain from being in a small class is **13.9** points.
    This is approximately 20% of the standard deviation in test scores,
    representing a moderate increase.

b.  **Prepare**:

$$
\left\{\begin{aligned}
    &H_0: \beta_{\text{SmallClass}} = 0 \\
    &H_1: \beta_{\text{SmallClass}} \neq 0
    \end{aligned}\right.
$$ Let the significance level be $\alpha = 0.05$.

**Calculate**:

$$
t^* = \frac{13.9 - 0}{2.5} \approx 5.56
$$ **Conclude**:

```         
Since $t^* > Z_{0.025} = 1.96$, we reject $H_0$. There is significant evidence that the estimated effect of class size on test scores is different from 0.
```

c.  Confidence Interval:

$$
\hat{\beta}_{\text{SmallClass}} \pm Z_{0.005} \cdot SE(\hat{\beta}_{\text{SmallClass}}) = 13.9 \pm 2.576 \times 2.5 = [7.46, 20.34]
$$

d.  Yes. Students were randomly assigned to small or regular classes, so
    that $\text{SmallClass}$ is independent of characteristics of the
    student, including those affecting $\text{TestScore}$, that is $u$.

## Computer Exercises

[(hprice.csv)](hprice.csv){target="_blank" rel="noopener"}

``` r
# remove pre-load data, and import package
rm(list = ls())
library(tidyverse)

# set directory and import data
house <- read.csv("data/hprice.csv")
```

```{r}
#| echo: false

rm(list = ls())
library(tidyverse)
library(here)

house <- read.csv(here("data", "hprice.csv")) 
```

### House Price: Summary Statistics

``` r
summary(house$price)
```

```{r}
#| echo: false        

summary(house$price)
```

``` r
hprice_ttest <- t.test(
  house$price,
  conf.level = 0.95
)
hprice_ttest$conf.int
```

```{r}
#| echo: false        

hprice_ttest <- t.test(
  house$price,
  conf.level = 0.95
)
hprice_ttest$conf.int
```

This confidence interval suggests that we are 95% confident that the
true mean house price in the population falls between \$269,006 and
\$307,609 (in \$1000 units). This means that, based on the sample data,
the average price of houses in Boston in 1990 is likely within this
range.

### House Price: Average Price

For the house built in colonial style:

``` r
house %>% 
  filter(colonial == 1) %>% 
  summarize(avg_hprice = mean(price))
```

```{r}
#| echo: false        

house %>% 
  filter(colonial == 1) %>% 
  summarize(avg_hprice = mean(price))
```

and for the house *not* built in colonial style:

``` r
house %>% 
  filter(colonial == 0) %>% 
  summarize(avg_hprice = mean(price))
```

```{r}
#| echo: false        

house %>% 
  filter(colonial == 0) %>% 
  summarize(avg_hprice = mean(price))
```

### House Price: Scatter Plot

``` r
ggplot(house, aes(x = sqrft, y = price)) +
  geom_point() +
  labs(
    x = "house size (sqrft)", 
    y = "house price (price)"
  )
```

```{r}
#| echo: false        

ggplot(house, aes(x = sqrft, y = price)) +
  geom_point() +
  labs(
    x = "house size (sqrft)", 
    y = "house price (price)"
  )
```

``` r
cor(house$price, house$sqrft)
```

```{r}
#| echo: false        

cor(house$price, house$sqrft)
```

Because $\text{Corr}(price, \textit{sqrft}) > 0.7$, the strength of the
relationship is strong. Also by inspecting the scatter plot, we can
conclude that fitting a linear model would be reasonable.

### House Price: Regression (1)

Given the following specification:

$$
\textit{price}_i = \beta_0 + \beta_1 \textit{sqrft}_i + u_i
$$ we propose the hypothesis:

$$
\left\{\begin{aligned}
&H_0: \beta_1 = 0 \\
&H_1: \beta_1 \neq 0
\end{aligned}\right.
$$

``` r
reg1 <- lm(
  formula = price ~ sqrft,
  data = house
)
summary(reg1)
```

```{r}
#| echo: false        

reg1 <- lm(
  formula = price ~ sqrft,
  data = house
)
summary(reg1)
```

From the regression result, we have the $p$-value of the coefficient of
house size ($\textit{sqrft}$) is very small ($<5\%$), hence we reject
the null hypothesis. There is significant evidence that the coefficient
on house size is different from zero.

### House Price: Regression (2)

``` r
reg2 <- lm(
  formula = price ~ colonial,
  data = house
)
reg2$coefficients
```

```{r}
#| echo: false        

reg2 <- lm(
  formula = price ~ colonial,
  data = house
)
reg2$coefficients
```

From the regression result, we have the estimated coefficients of
$\textit{colonial}$ is same as colonial gap, which is the gap between
two means in question (b).
