---
title: 'Homework 3 Answers'
fig-cap: false
editor: 
  markdown: 
    wrap: 72
---

-   If you find any typos or errors, please feel free to contact me via
    email at r13323002\@ntu.edu.tw. I would appreciate it and will
    correct the solution to prevent any misunderstandings.

-   For some R commands marked with ‘OPTIONAL,’ you don’t need to learn
    them. You will be completely fine on the computer quiz even if you
    choose to ignore them.

-   If you’re having trouble understanding the code, don’t hesitate to
    ask me for help. I’ll do my best to assist you during office hours
    or after the TA session.

## Textbook Exercises

### Exercise 6.1

Denote $\bar{R}_{(j)}^2$ as the adjusted R-squared of estimation in
column $j \in \{1,2,3\}$. Note that

$$
1-R^2 = \frac{SSR}{TSS}
$$

we can compute:

$$
\bar{R}^2_{(1)}
= 1 - \frac{n - 1}{n - k - 1} \frac{SSR}{TSS}
= 1 - \frac{7178 - 1}{7178 - 2 - 1} (1 - 0.165)
\approx 0.1648
$$

$$
\bar{R}^2_{(2)}
= 1 - \frac{n - 1}{n - k - 1} \frac{SSR}{TSS}
= 1 - \frac{7178 - 1}{7178 - 3 - 1} (1 - 0.182) \approx 0.1817
$$

$$
\bar{R}^2_{(3)}
= 1 - \frac{n - 1}{n - k - 1} \frac{SSR}{TSS}
= 1 - \frac{7178 - 1}{7178 - 6 - 1} (1 - 0.185) \approx 0.1843
$$

### Exercise 6.2

a.  On average, workers with college degrees earn \$10.47/hour more than
    workers with only high school degrees.\
b.  On average, men earn \$4.69/hour more than women.

### Exercise 6.3

a.  Other things being equal, a worker earns \$0.61/hour more for each
    additional year of age on average.

b.  Sally’s earnings prediction:

$$
    \begin{aligned}
    \widehat{AHE}
    &= \hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 + \hat{\beta}_3 X_3 \\
    &= 0.11 + 10.44 \cdot 1 - 4.56 \cdot 1 + 0.61 \cdot 29 = 23.68 \text{ (dollars per hour)}
    \end{aligned}
$$

Betsy’s earnings prediction:

$$
    \begin{aligned}
    \hat{AHE}
    &= \hat{\beta}_0 + \hat{\beta}_1 X_1 + \hat{\beta}_2 X_2 + \hat{\beta}_3 X_3 \\
    &= 0.11 + 10.44 \cdot 1 - 4.56 \cdot 1 + 0.61 \cdot 34 = 26.73 \text{ (dollars per hour)}
    \end{aligned}
$$

### Exercise 6.4

a.  

-   Other things being equal, workers in the Northeast earn \$0.74/hour
    more than workers in the West on average.

-   Other things being equal, workers in the Midwest earn \$1.54/hour
    less than workers in the West on average.

-   Other things being equal, workers in the South earn \$0.44/hour less
    than workers in the West on average.

b.  To avoid perfect multicollinearity, the variable $\textit{West}$ is
    omitted. If $\textit{West}$ were included, the intercept could be
    written as a perfect linear function of the four regional
    regressors.

c.  The expected difference in earnings between Juanita and Jennifer is:

$$
    \hat{\beta}_6 - \hat{\beta}_5 = -0.44 - (-1.54) = 1.1 \text{ (dollars per hour)}
$$

### Exercise 7.2

a.  **Prepare**:

$$
    \left\{\begin{aligned}
    &H_0: \beta_1 = 0 \\
    &H_1: \beta_1 \neq 0
    \end{aligned}\right.
$$

Let the significance level be 5%.

**Calculate**: The 95% confidence interval for $\beta_1$ is

$$
    \hat{\beta}_1 \pm Z_{0.025} \cdot SE(\hat{\beta}_1) = 0.352 \pm 1.96 \times 0.021 = [0.3108, 0.3932]
$$

**Conclusion**:

Since 0 is not in the 95% confidence interval, we reject $H_0$. There is
significant evidence that high school earnings differences exist.

b.  **Prepare**:

$$
    \left\{\begin{aligned}
    &H_0: \beta_2 = 0 \\
    &H_1: \beta_2 \neq 0
    \end{aligned}\right.
$$

Let the significance level be 5%.

**Calculate**: The 95% confidence interval for $\beta_2$ is

$$
    \hat{\beta}_2 \pm Z_{0.025} \cdot SE(\hat{\beta}_2) = 0.458 \pm 1.96 \times 0.021 = [0.4168, 0.4992]
$$

Since 0 is not in the 95% confidence interval, we reject $H_0$. There is
significant evidence that male-female earnings differences exist.

### Exercise 7.3

a.  **Prepare**:

$$
    \left\{\begin{aligned}
    &H_0: \beta_3 = 0 \\
    &H_1: \beta_3 \neq 0
    \end{aligned}\right.
$$

Let the significance level be 5%.

**Calculate**: The 95% confidence interval for $\beta_3$ is

$$
    \hat{\beta}_3 \pm Z_{0.025} \cdot SE(\hat{\beta}_3) = 0.011 \pm 1.96 \times 0.001 = [0.0090, 0.0130]
$$

Since 0 is not in the 95% confidence interval, we reject $H_0$. There is
significant evidence that age is an important determinant of earnings.

b.  95% Confidence Interval for the expected earnings difference:

$$
\begin{aligned}
\Delta X_3 \times \left(\hat{\beta}_3 \pm Z_{0.025} \cdot SE(\hat{\beta}_3)\right)
= (40 - 30) \times (0.011 \pm 1.96 \times 0.001) = [0.0904, 0.1296]
\end{aligned}
$$

### Exercise 7.4

a.  To perform an appropriate hypothesis test on regional differences,
    we conduct the following procedures step-by-step:

<!-- -->

1.  Claim the null hypothesis:

$$
\left\{\begin{aligned}
&H_0: \beta_4 = \beta_5 = \beta_7 = 0 \\
&H_1: \text{Not all are zero}
\end{aligned}\right.
$$

In other words, we are testing if the difference between *restricted
model* and *unrestricted model* is statistically significant.

-   Restricted Model ($\beta_4 = \beta_5 = \beta_7 = 0$)

$$
\text{AWE} = \beta_0 + \sum_{i=1}^3 \beta_i X_i + u
$$

-   Unrestricted Model

$$
\text{AWE} = \beta_0 + \sum_{i=1}^3 \beta_i X_i + \beta_4 X_4 + \beta_5 X_5 + \beta_6 X_6 + v
$$

2.  Next, we derive the pivotal quantity:

$$
\begin{aligned}
\phi &= \frac{(SSE_r - SSE_u) / q}{(SSE_u) / (n - k_u - 1)} \sim F(q, n - k_u - 1)
\end{aligned}
$$

where $SSE_r$ is the SSE of the restricted model, and $SSE_u$ is the SSE
of the unrestricted model. Besides, $n$ is the number of samples, $k_u$
is the number of independent variables used in the unrestricted model,
and $q$ is the number of restricted variables.

In this case, $n=10,000$, $k_u=6$ and $q=3$, therefore,

$$
\phi \sim F(3, 10000-6-1) \sim F(3, 9993)
$$

3.  Given $\alpha = 0.05$, we can construct the rejection region (RR):

$$
RR = \{\phi \mid \phi > F_{0.05}(3, 9993) = 2.6058\}
$$

4.  Finally, since

$$
\phi^* = 21.87 \in RR
$$

we reject $H_0$. There is significant evidence that regional differences
exist.

b.  

    i.  Consider the model in column (3). The expected difference
        between Juan and Mel is $\hat{\beta}_4$, hence the 95%
        confidence interval for the expected difference between Juan and
        Mel is:

$$
\hat{\beta}_4 \pm Z_{0.025} \cdot SE(\hat{\beta}_4) = 0.175 \pm 1.96 \times 0.037 = [0.1025, 0.2475]
$$

ii. The expected difference between Juan and Ari is:

$$
\beta_4 \times (X_{4,\text{Juan}} - X_{4,\text{Ari}}) + \beta_7 \times (X_{7,\text{Juan}} - X_{7,\text{Ari}}) = \beta_4 - \beta_7
$$

A 95% confidence interval could be constructed using the general
methods. In this case, an easy way to do this is to omit East from the
regression and replace it with $X_6 = \textit{West}$. In this new
regression, the coefficient on North measures the difference in wages
between the North and East, and a 95% confidence interval can be
computed directly. 

## Computer Exercises

[(hprice.csv)](hprice.csv){target="_blank" rel="noopener"}

[(wage.csv)](wage.csv){target="_blank" rel="noopener"}

``` r
# remove pre-load data, and import package
rm(list = ls())
library(tidyverse)
library(estimatr) # for robust SE
library(texreg) # compile regression result into one table
library(car) # short for "Companion to Applied Regression", used for hypothesis testing

house <- read.csv("data/hprice.csv")
wage <- read.csv("data/wage.csv")
```

```{r}
#| echo: false
# remove pre-load data, and import package
rm(list = ls())
library(tidyverse)
library(estimatr) # for robust SE
library(texreg) # compile regression result into one table
library(car) # short for "Companion to Applied Regression", used for hypothesis testing
library(here)


house <- read.csv(here("data", "hprice.csv")) 
wage <- read.csv(here("data", "wage.csv")) 
```

### Question 1: House Price

a.  See the correlation matrix below:

``` r
house %>%
  select(price, sqrft, bdrms, lotsize) %>%
  cor()
```

```{r}
#| echo: false

house %>%
  select(price, sqrft, bdrms, lotsize) %>%
  cor()
```

b.  See the estimation results below:

``` r
reg_1b <- lm_robust(
  formula = price ~ sqrft + bdrms,
  data = house,
  se_type = "HC1"
)
summary(reg_1b)
```

```{r}
#| echo: false

reg_1b <- lm_robust(
  formula = price ~ sqrft + bdrms,
  data = house,
  se_type = "HC1"
)
summary(reg_1b)
```

c.  Other things being equal, the estimated price increase in
    $14.4160553$ thousand for a house with one more bedroom on average.

d.  $62.8307451\%$ of the variation in price is explained by square
    footage and the number of bedrooms.

e.  Definitely. Take variable $\textit{bdrms}$ for example, we can
    easily tell the following relationship in common sense:

-   $\text{Cov}(\textit{price}, \textit{bdrms}) > 0$

-   $\text{Cov}(\textit{sqrt}, \textit{bdrms}) > 0$

as $\textit{bdrms}$ is omitted from the model, we would expect an upward
bias (overestimates) in the estimated coefficient.

``` r
reg_1e <- lm_robust(
  formula = price ~ sqrft,
  data = house
)

# build regression table
screenreg(
  list(reg_1b, reg_1e),
  digits = 4
)
```

```{r}
#| echo: false

reg_1e <- lm_robust(
  formula = price ~ sqrft,
  data = house
)

# build regression table
screenreg(
  list(reg_1b, reg_1e),
  digits = 4
)
```

The regression result matches our interpretation, that is, omitting
variable will lead to overestimates (0.1393 \> 0.1281).

f.  From the correlation matrix in (a), we have
    $\text{Corr}(\textit{lotsize}, \textit{sqrft}) \neq 1$, so there
    will be no perfect multicollinearity if we add $\textit{lotsize}$
    into the regression in question (b).

``` r
reg_1f <- lm_robust(
  formula = price ~ sqrft + bdrms + lotsize,
  data = house,
  se_type = "HC1"
)
summary(reg_1f)
```

```{r}
#| echo: false

reg_1f <- lm_robust(
  formula = price ~ sqrft + bdrms + lotsize,
  data = house,
  se_type = "HC1"
)
summary(reg_1f)
```

Other things being equal, the estimated price increase in $0.0021293$
thousand for a house with one more square feet about the size of the lot
on average.

### Question 2: Wage

a.  According to the classical human capital theory, we would expect all
    the signs to be positive. See the estimation results below:

``` r
reg_2a <- lm_robust(
  formula = wage ~ exper + stotal + jc + univ,
  data = wage,
  se_type = "HC1"
)
summary(reg_2a)
```

```{r}
#| echo: false

reg_2a <- lm_robust(
  formula = wage ~ exper + stotal + jc + univ,
  data = wage,
  se_type = "HC1"
)
summary(reg_2a)
```

The estimates are all positive, which match our intuition. The estimated
hourly wages increase in $0.4093446$ with 10 more work experience on
average. On average, the hourly wage for a worker with no experience, no
college degree of any sort and with 0 standardized score is $4.0299005$.

b.  We want to test the following hypothesis:

$$
\begin{aligned}
&H_0: \beta_3 = \beta_4 = 0 \\
&H_1: \text{Not all are zero}
\end{aligned}
$$

See the testing result below:

``` r
linearHypothesis(
  reg_2a,
  c("jc = 0", "univ = 0"),
  test = "F"
)
```

```{r}
#| echo: false

linearHypothesis(
  reg_2a,
  c("jc = 0", "univ = 0"),
  test = "F"
)
```

From the testing result, we have its $p$-value is very small ($<5\%$).
There is significant evidence that the effects of attending two-year or
four-year college are different.

c.  See the estimation results below:

``` r
reg_2c <- lm_robust(
  formula = wage ~ exper + stotal + jc + univ + AA + BA,
  data = wage,
  se_type = "HC1"
)
summary(reg_2c)
```

```{r}
#| echo: false

reg_2c <- lm_robust(
  formula = wage ~ exper + stotal + jc + univ + AA + BA,
  data = wage,
  se_type = "HC1"
)
summary(reg_2c)
```

d.  

    i.  **Individual significance**

        From the preceding estimation result, we have both $\textit{AA}$
        and $\textit{BA}$ not individually significant under $5\%$
        significance level because both $p$-values are greater than
        0.05.

    ii. **Joint significance**

        ``` r
        linearHypothesis(
          reg_2c,
          c("AA = 0", "BA = 0"),
          test = "F"
        )
        ```

        Since $p$-value is greater than $0.05$, we can't reject the null
        hypothesis that both coefficients are equal to $0$. There is no
        evidence that the effects of college degree exist.

    iii. **Change in** $R^2$

    ```         
      By comparing the estimation results in question (a), (c), we have the
      adjusted $R^2$ from specification (1) is greater than (2). In other
      words, adding these two variables doesn't improve explaining the
      variation in wage.

      In conclusion, we suggest that the two dummy variables $AA$ and $BA$
      should not be kept in the regression specification.
    ```
