---
title: 'Homework 4 Answers'
fig-cap: false
editor:
  markdown:
    wrap: 72
---

-   If you find any typos or errors, please feel free to contact me via
    email at r13323002\@ntu.edu.tw. I would appreciate it and will
    correct the solution to prevent any misunderstandings.

-   For some R commands marked with ‘OPTIONAL,’ you don’t need to learn
    them. You will be completely fine on the computer quiz even if you
    choose to ignore them.

-   If you’re having trouble understanding the code, don’t hesitate to
    ask me for help. I’ll do my best to assist you during office hours
    or after the TA session.

## Textbook Exercises

### Exercise 8.2

a.  The expected change in price of building a 1500-square-foot addition
    to a house is:


$$
1500 \times (0.00042 \times 100\%) = 63\%
$$

and the 99% confidence interval for the percentage change in price:


$$
1500 \times (0.00042 \pm 2.576 \times 0.000038) \times 100\% = [48.32\%, 77.68\%]
$$

b.  The coefficient in a log-log regression is interpreted as the
    percentage increase in the dependent variable with a $1\%$
    increase in the independent variable.

    Hence a 1% change in $\textit{size}$ is associated with a
    $0.69\%$ change in $\textit{price}$; when the house doubles in
    size, the house price is expected to increase by:


$$
0.69\% \times 100 = 69\%
$$

c.  The house price is expected to increase by:


$$
0.027 \times 100\% = 2.7\%
$$

**Prepare**:


$$
\left\{\begin{aligned} &H_0: \beta_{\textit{view}} = 0 \\ &H_1: \beta_{\textit{view}} \neq 0 \end{aligned}\right.
$$

Let the significance level be $\alpha = 0.01$.

**Calculate**: the $99\%$ confidence interval for $\beta_{\text{view}}$:


$$
\hat{\beta}_{\text{view}} \pm Z_{0.005} \times SE(\hat{\beta}_{\text{view}}) = 0.027 \pm 2.576 \times 0.028 = [-4.51\%, 9.91\%]
$$

**Conclusion**:

Since 0 is in the confidence interval, we do **not** reject $H_0$.
here is no evidence that view has a significant effect on price.

d.  The expected change in price for adding two bedrooms:


$$
2 \times (0.0036 \times 100\%) = 0.72\%
$$

**Prepare**:


$$
\left\{\begin{aligned} &H_0: \beta_{\textit{bedroom}} = 0 \\ &H_1: \beta_{\textit{bedroom}} \neq 0 \end{aligned}\right.
$$

Let the significance level be $\alpha = 0.01$.

**Calculate**:


$$
t^* = \frac{0.0036}{0.037} \approx 0.0973
$$

**Conclusion**:

Since $t^* < Z_{0.005} = 2.576$, we do not reject $H_0$.
There is no evidence that bedrooms have a significant effect on price. Because size and bedrooms are likely correlated, assessing the relative significance of each is difficult.

e.  **Prepare**:


$$
\left\{\begin{aligned} &H_0: \beta_{\textit{condition}} = 0 \\ &H_1: \beta_{\textit{condition}} \neq 0 \end{aligned}\right.
$$

Let the significance level be $\alpha = 0.01$.

**Calculate**:


$$
t^* = \frac{0.12}{0.036} \approx 3.3333
$$

**Conclusion**:

Since $t^* > Z_{0.005} = 2.576$, we reject $H_0$. There is significant evidence that condition affects house prices.

f.  **Prepare**:


$$
\left\{\begin{aligned} &H_0: \beta_{\textit{pool} \times \textit{view}} = 0 \\ &H_1: \beta_{\textit{pool} \times \textit{view}} \neq 0 \end{aligned}\right.
$$

Let the significance level be $\alpha = 0.01$.

**Calculate**:


$$
t^* = \frac{0.0022}{0.10} = 0.022
$$

**Conclusion**:

Since $t^* < Z_{0.005} = 2.576$, we do not reject $H_0$. There is no evidence that the interaction effect is significant.

Now, when other thing being equal, the change of expected price when adding a view to a house with a pool is:


$$
(0.027 + 0.0022) \times 100\% = 2.92\%
$$

while for a house without a pool is:


$$
0.027 \times 100\% = 2.7\%
$$

### Exercise 8.7

a.  i.  On average, $\ln(\text{Earnings})$ for women is 0.44 lower
        than for men.\

$$
E[\ln(\text{Earnings}) | \text{Female} = 0] - E[\ln(\text{Earnings}) | \text{Female} = 1] = 0.44
$$

This implies:\

$$
\ln \left( \frac{\text{Men's avg earnings}}{\text{Women's avg earnings}} \right) = 0.44
$$


$$
\frac{\text{Men's avg earnings}}{\text{Women's avg earnings}} = e^{0.44} \approx 1.553
$$
 Thus, on average, men earn 55.3% more than women.

    ii. The error term has a standard deviation of 2.65 (measured in
        log-points).

    iii. Yes, but the regression does not control for firm size,
         industry, profitability, experience, etc.

    iv. No, because gender discrimination means that identical workers
        (except for gender) are paid different wages. Controlling for
        characteristics affecting productivity (e.g., education,
        experience) is necessary.

b.  i.  Holding other factors constant, a 1% increase in Market Value is
        associated with a 0.37% increase in earnings.

    ii. Female is correlated with Market Value and Return, which are
        important for explaining $\ln(\text{Earnings})$. Thus, the
        regression in (a) suffers from omitted variable bias.

c.  No, the coefficients are not stable. Ignoring the effect of Return,
    the omitted variable bias formula suggests that Female is negatively
    correlated with $\ln(\text{Market Value})$.

------------------------------------------------------------------------

### Exercise 9.6

a.  First note the following equalities:


$$
\bar{Y}_{\text{new}} = \frac{\sum_{i=1}^{100} Y_i}{100} = \frac{2 \sum_{i=1}^{50} Y_i}{100} = \bar{Y}_{\text{old}}
$$


$$
\bar{X}_{\text{new}} = \frac{\sum_{i=1}^{100} X_i}{100} = \frac{2 \sum_{i=1}^{50} X_i}{100} = \bar{X}_{\text{old}}
$$

and by the formula of $\hat{\beta}_1$, we have:


$$
\begin{aligned} \hat{\beta}_1^{new} &= \frac{\sum_{i=1}^{100} (X_i - \overline{X}_{new}) (Y_i - \overline{Y}_{new})}{\sum_{i=1}^{100} (X_i - \overline{X}_{new})^2} \\ &= \frac{2\sum_{i=1}^{50} (X_i - \overline{X}_{old}) (Y_i - \overline{Y}_{old})}{2\sum_{i=1}^{50} (X_i - \overline{X}_{old})^2} = \hat{\beta}_1^{old} = 73.9 \end{aligned}
$$

and it's clear that $\hat{\beta}_0^{new} = \hat{\beta}_0^{old}$ since:


$$
\begin{aligned} \hat{\beta}_0^{new} &= \overline{Y}_{new} - \hat{\beta}_1^{new} \overline{X}_{new} \\ &= \overline{Y}_{old} - \hat{\beta}_1^{old} \overline{X}_{old} = \hat{\beta}_0^{old} = 49.2 \end{aligned}
$$

Moreover,


$$
\begin{aligned} SSR_{old} &= (50 - 2) \cdot SER_{old}^2 \\ &= (50 - 2) \cdot 13.4^2 = 8618.88 \end{aligned}
$$


$$
\begin{aligned} SSR_{new} &= 2 \times SSR_{old} \\ &= 2 \times 8618.88 = 17237.76 \end{aligned}
$$


$$
\begin{aligned} SER_{new} &= \sqrt{\frac{1}{100 - 2} \cdot SSR_{new}} \\ &= \sqrt{\frac{1}{100 - 2} \cdot 17237.76} \approx 13.2626 \end{aligned}
$$


$$
\begin{aligned} R^2_{new} &= \frac{\sum_{i=1}^{100} (\hat{Y}_i^{new} - \overline{Y}_{new})^2}{\sum_{i=1}^{100} (Y_i^{new} - \overline{Y}_{new})^2} \\ &= \frac{2 \sum_{i=1}^{50} (\hat{Y}_i^{old} - \overline{Y}_{old})^2}{2 \sum_{i=1}^{50} (Y_i^{old} - \overline{Y}_{old})^2} = R^2_{old} = 0.78. \end{aligned}
$$

Besides,


$$
\begin{aligned} SE\left(\hat{\beta}_1^{new}\right) &= \sqrt{ \frac{1}{100} \times \frac{ \frac{1}{100-2} \sum_{i=1}^{100} (X_i - \overline{X}_{new})^2 \hat{u}_{i,new}^2 }{ \left[\frac{1}{100} \sum_{i=1}^{100} (X_i - \overline{X}_{new})^2\right]^2 } } \\ &= \sqrt{ \frac{1}{100} \times \frac{ \frac{2}{100-2} \sum_{i=1}^{50} (X_i - \overline{X}_{old})^2 \hat{u}_{i,old}^2 }{ \left[\frac{2}{100} \sum_{i=1}^{50} (X_i - \overline{X}_{old})^2\right]^2 } } \\ &= \sqrt{ \hat{\sigma}_{\beta_1,old}^2 \times \frac{50-2}{100-2} } \\ &= \sqrt{ 16.4^2 \times \frac{50-2}{100-2} } \approx 11.4776 \end{aligned}
$$


$$
\begin{aligned} SE\left(\hat{\beta}_0^{new}\right) &= \sqrt{ \frac{1}{100} \times \frac{ \frac{1}{100-2} \sum_{i=1}^{100} \left(1 - \frac{\overline{X}_{new}}{\sum_{i=1}^{100} X_i^2}X_i\right) \hat{u}_{i,new}^2 }{ \left[\frac{1}{100} \sum_{i=1}^{100} \left(1 - \frac{\overline{X}_{new}}{\sum_{i=1}^{100} X_i^2}X_i\right)\right]^2 } } \\ &= \sqrt{ \frac{1}{100} \times \frac{ \frac{2}{100-2} \sum_{i=1}^{50} \left(1 - \frac{\overline{X}_{old}}{2\sum_{i=1}^{50} X_i^2}X_i\right) \hat{u}_{i,old}^2 }{ \left[\frac{2}{100} \sum_{i=1}^{50} \left(1 - \frac{\overline{X}_{old}}{2\sum_{i=1}^{50} X_i^2}X_i\right)\right]^2 } } \\ &\approx \sqrt{ \hat{\sigma}_{\hat{\beta}_0,old}^2 \times \frac{50-2}{100-2} } \\ &= \sqrt{ 23.5^2 \times \frac{50-2}{100-2} } \approx 16.4466 \end{aligned}
$$

In conclusion, we can determin the regression statistics:


$$
\hat{Y} = \underset{(16.4466)}{49.2} + \underset{11.4776}{73.9} X, \quad SER = 13.2626, \quad R^2 = 0.78
$$

b.  Because half of the observations are identical to the other half,
    the data are not independent. Thus, the observations are not i.i.d..

## Computer Exercises

[(cps98.csv)](cps98.csv){target="_blank" rel="noopener"}

``` r
# remove pre-load data, and import package
rm(list = ls())
library(tidyverse)
library(estimatr) # for robust SE
library(texreg) # compile regression result into one table
library(car) # short for "Companion to Applied Regression", used for hypothesis testing

cps <- read.csv("data/cps98.csv")
```

```{r}
#| echo: false
# remove pre-load data, and import package
rm(list = ls())
library(tidyverse)
library(estimatr) # for robust SE
library(texreg) # compile regression result into one table
library(car) # short for "Companion to Applied Regression", used for hypothesis testing
library(here)

cps <- read.csv(here("data", "cps98.csv"))
```

### Wage: Model 1

See the estimation results below:

``` r
reg_1a <- lm_robust(
  formula = ahe ~ age + female + bachelor,
  data = cps,
  se_type = "HC1"
)
summary(reg_1a)
```

```{r}
#| echo: false

reg_1a <- lm_robust(
  formula = ahe ~ age + female + bachelor,
  data = cps,
  se_type = "HC1"
)
summary(reg_1a)
```

For the prediction of change in earning when age increase from 25 to 26:

``` r
# construct two groups of sample, which differ only in age
group_age25 <- data.frame(
  age = 25,
  female = 0,
  bachelor = 0
)
group_age26 <- data.frame(
  age = 26,
  female = 0,
  bachelor = 0
)

# predict the wage for these groups, and calculate the wage difference
predict(reg_1a, group_age26) - predict(reg_1a, group_age25)
```

```{r}
#| echo: false

# construct two groups of sample, which differ only in age
group_age25 <- data.frame(
  age = 25,
  female = 0,
  bachelor = 0
)
group_age26 <- data.frame(
  age = 26,
  female = 0,
  bachelor = 0
)

# predict the wage for these groups, and calculate the wage difference
predict(reg_1a, group_age26) - predict(reg_1a, group_age25)
```

For the prediction of change in earning when age increase from 33 to 34:

``` r
# construct two groups of sample, which differ only in age
group_age33 <- data.frame(
  age = 33,
  female = 0,
  bachelor = 0
)
group_age34 <- data.frame(
  age = 34,
  female = 0,
  bachelor = 0
)

# predict the wage for these groups, and calculate the wage difference
predict(reg_1a, group_age26) - predict(reg_1a, group_age25)
```

```{r}
#| echo: false

# construct two groups of sample, which differ only in age
group_age33 <- data.frame(
  age = 33,
  female = 0,
  bachelor = 0
)
group_age34 <- data.frame(
  age = 34,
  female = 0,
  bachelor = 0
)

# predict the wage for these groups, and calculate the wage difference
predict(reg_1a, group_age26) - predict(reg_1a, group_age25)
```

Therefore, if age increases from 25 to 26, hourly earnings are predicted
to increase by $0.314; while if age increases from 33 to 34, hourly
earnings are predicted to increase by$0.314.

### Wage: Model 2

``` r
reg_1b <- lm_robust(
  formula = log(ahe) ~ age + female + bachelor,
  data = cps,
  se_type = "HC1"
)
summary(reg_1b)
```

```{r}
#| echo: false

reg_1b <- lm_robust(
  formula = log(ahe) ~ age + female + bachelor,
  data = cps,
  se_type = "HC1"
)
summary(reg_1b)
```

``` r
predict(reg_1b, group_age26) - predict(reg_1b, group_age25)
```

```{r}
#| echo: false

predict(reg_1b, group_age26) - predict(reg_1b, group_age25)
```

``` r
predict(reg_1b, group_age34) - predict(reg_1b, group_age33)
```

```{r}
#| echo: false

predict(reg_1b, group_age34) - predict(reg_1b, group_age33)
```

Therefore, if age increases from 25 to 26, hourly earnings are predicted
to increase by 2.142%; while if age increases from 33 to 34, hourly
earnings are predicted to increase by 2.142%.

### Wage: Model 3

``` r
reg_1c <- lm_robust(
  formula = log(ahe) ~ log(age) + female + bachelor,
  data = cps,
  se_type = "HC1"
)
summary(reg_1c)
```

```{r}
#| echo: false

reg_1c <- lm_robust(
  formula = log(ahe) ~ log(age) + female + bachelor,
  data = cps,
  se_type = "HC1"
)
summary(reg_1c)
```

``` r
predict(reg_1c, group_age26) - predict(reg_1c, group_age25)
```

```{r}
#| echo: false

predict(reg_1c, group_age26) - predict(reg_1c, group_age25)
```

``` r
predict(reg_1c, group_age34) - predict(reg_1c, group_age33)
```

```{r}
#| echo: false

predict(reg_1c, group_age34) - predict(reg_1c, group_age33)
```

Therefore, if age increases from 25 to 26, hourly earnings are predicted
to increase by 2.498%; while if age increases from 33 to 34, hourly
earnings are predicted to increase by 1.901%.

### Wage: Model 4

``` r
reg_1d <- lm_robust(
  formula = log(ahe) ~ age + I(age^2) + female + bachelor,
  data = cps,
  se_type = "HC1"
)
summary(reg_1d)
```

```{r}
#| echo: false

reg_1d <- lm_robust(
  formula = log(ahe) ~ age + I(age^2) + female + bachelor,
  data = cps,
  se_type = "HC1"
)
summary(reg_1d)
```

``` r
predict(reg_1d, group_age26) - predict(reg_1d, group_age25)
```

```{r}
#| echo: false

predict(reg_1d, group_age26) - predict(reg_1d, group_age25)
```

``` r
predict(reg_1d, group_age34) - predict(reg_1d, group_age33)
```

```{r}
#| echo: false

predict(reg_1d, group_age34) - predict(reg_1d, group_age33)
```

Therefore, if age increases from 25 to 26, hourly earnings are predicted
to increase by 4.05%; while if age increases from 33 to 34, hourly
earnings are predicted to increase by 0.338%.

### Wage: Model 5 (woman)

``` r
reg_1e <- lm_robust(
  formula = log(ahe) ~ age + I(age^2) + female + bachelor + female*bachelor,
  data = cps,
  se_type = "HC1"
)
summary(reg_1e)
```

```{r}
#| echo: false

reg_1e <- lm_robust(
  formula = log(ahe) ~ age + I(age^2) + female + bachelor + female*bachelor,
  data = cps,
  se_type = "HC1"
)
summary(reg_1e)
```

``` r
fem_age30_ba <- data.frame(
  age = 30,
  female = 1,
  bachelor = 1
)
predict(reg_1e, fem_age30_ba)
```

```{r}
#| echo: false

fem_age30_ba <- data.frame(
  age = 30,
  female = 1,
  bachelor = 1
)
predict(reg_1e, fem_age30_ba)
```

``` r
fem_age30 <- data.frame(
  age = 30,
  female = 1,
  bachelor = 0
)
predict(reg_1e, fem_age30)
```

```{r}
#| echo: false

fem_age30 <- data.frame(
  age = 30,
  female = 1,
  bachelor = 0
)
predict(reg_1e, fem_age30)
```

``` r
predict(reg_1e, fem_age30_ba) - predict(reg_1e, fem_age30)
```

```{r}
#| echo: false

predict(reg_1e, fem_age30_ba) - predict(reg_1e, fem_age30)
```

According to the estimation results, we can answer the following
question:

-   **Coefficient on interaction term**: The extra effect of
    $\textit{bachelor}$ on $\ln(\textit{ATE})$ for woman relative
    to man.

-   $\widehat{\ln(\textit{ATE})}$ **for 30-year-old woman with a
    bachelor's degree**: $2.677$.

-   $\widehat{\ln(\textit{ATE})}$ **for 30-year-old woman without a
    bachelor's degree**: $2.246$.

-   **Predicted difference**: $0.431$.

### Wage: Model 5 (man)

``` r
male_age30_ba <- data.frame(
  age = 30,
  female = 0,
  bachelor = 1
)
predict(reg_1e, male_age30_ba)
```

```{r}
#| echo: false

male_age30_ba <- data.frame(
  age = 30,
  female = 0,
  bachelor = 1
)
predict(reg_1e, male_age30_ba)
```

``` r
male_age30 <- data.frame(
  age = 30,
  female = 0,
  bachelor = 0
)
predict(reg_1e, male_age30)
```

```{r}
#| echo: false

male_age30 <- data.frame(
  age = 30,
  female = 0,
  bachelor = 0
)
predict(reg_1e, male_age30)
```

``` r
predict(reg_1e, male_age30_ba) - predict(reg_1e, male_age30)
```

```{r}
#| echo: false

predict(reg_1e, male_age30_ba) - predict(reg_1e, male_age30)
```

-   $\widehat{\ln(\textit{ATE})}$ **for 30-year-old man with a
    bachelor's degree**: $2.810$.

-   $\widehat{\ln(\textit{ATE})}$ **for 30-year-old man without a
    bachelor's degree**: $2.464$.

-   **Predicted difference**: $0.346$.

### Wage: Model 6 (testing)

We want to test the following hypothesis:

$$
\left\{\begin{aligned} &H_0: \beta_{\textit{age} \times \textit{bachelor}} = \beta_{\textit{age}^2 \times \textit{bachelor}} = 0 \\ &H_1: \text{Not all are zero} \end{aligned}\right.
$$

See the testing result below:

``` r
reg_1g <- lm_robust(
  formula = log(ahe) ~ age + I(age^2) + female + bachelor
    + female*bachelor + age*bachelor + I(age^2)*bachelor,
  data = cps,
  se_type = "HC1"
)
linearHypothesis(
  reg_1g,
  c("age:bachelor = 0", "I(age^2):bachelor = 0"),
  test = "F"
)
```

```{r}
#| echo: false

reg_1g <- lm_robust(
  formula = log(ahe) ~ age + I(age^2) + female + bachelor
    + female*bachelor + age*bachelor + I(age^2)*bachelor,
  data = cps,
  se_type = "HC1"
)
linearHypothesis(
  reg_1g,
  c("age:bachelor = 0", "I(age^2):bachelor = 0"),
  test = "F"
)
```

Because $p$-value is greater than $0.05$, we can't reject the null
hypothesis. There is no evidence that the effect of age on earnings is
different between high school graduates and college graduates.
