---
title: 'Homework 5 Answers'
fig-cap: false
editor:
  markdown:
    wrap: 72
---

-   If you find any typos or errors, please feel free to contact me via
    email at r13323002\@ntu.edu.tw. I would appreciate it and will
    correct the solution to prevent any misunderstandings.

-   For some R commands marked with ‘OPTIONAL,’ you don’t need to learn
    them. You will be completely fine on the computer quiz even if you
    choose to ignore them.

-   If you’re having trouble understanding the code, don’t hesitate to
    ask me for help. I’ll do my best to assist you during office hours
    or after the TA session.

## Textbook Exercises

### Exercise 10.1

a.  Other things being equal, if the tax on a case of beer increases by
$2, the number of lives saved is:

$$
2 \times 0.69 \times 885 = 1221.3
$$

and the 99% confidence interval for the number of lives saved is:

$$
[0.69 \pm Z_{0.005} \times 0.35] \times 2 \times 885 = [-374.53, 2817.13]
$$

b.  If New Jersey lowers its drinking age to 19, the predicted decrease
in traffic fatalities is:

$$
0.08 \times 885 = 70.8
$$

and the 95% confidence interval for the number of traffic fatalities
is:

$$
[0.08 \pm Z_{0.025} \times 0.07] \times 885 = [-50.62, 192.22]
$$

c.  If real income per capita in New Jersey increases by 3%, the
predicted increase in traffic fatalities is:

$$
3 \times 0.01 \times 1.79 \times 885 = 47.52
$$

and the 95% confidence interval for the number of traffic fatalities
is:

$$
[1.79 \pm Z_{0.025} \times 0.64] \times 0.01 \times 3 \times 885 = [14.22, 80.83]
$$

d.  Standard errors should be clustered into groups within which the
values of the dependent variable are likely to be correlated with
one another. In this case, the natural candidate group is to cluster
by state.

e.  If we believe that drinking age affects the number of fatalities due
to alcohol through a linear function, then age should be entered as
a continuous variable. If we allow for nonlinearities in this
relationship, we should allow for separate intercepts for each age.

### Exercise 10.5

$$
\begin{aligned}
Y_{11} &= \beta_1 X_{11} + \alpha_1 + \lambda_1 + u_{11} \\
&= \beta_0 + \beta_1 X_{11} + u_{11} \\
Y_{i1} &= \beta_1 X_{i1} + \alpha_i + \lambda_1 + u_{i1} \\
&= \beta_0 + \beta_1 X_{1i} + \gamma_i + u_{i1}, \quad i = 2, \dots, n \\
Y_{1t} &= \beta_1 X_{1t} + \alpha_1 + \lambda_t + u_{1t} \\
&= \beta_0 + \beta_1 X_{1t} + \delta_t + u_{1t}, \quad t = 2, \dots, T
\end{aligned}
$$

We have

$$
\begin{aligned}
\beta_0 &= \alpha_1 + \lambda_1 \\
\gamma_i &= \alpha_i + \lambda_1 - \beta_0 \\
&= \alpha_i - \alpha_1, \quad i = 2, \dots, n \\
\delta_t &= \alpha_1 + \lambda_t - \beta_0 \\
&= \lambda_t - \lambda_1, \quad t = 2, \dots, T.
\end{aligned}
$$

### Exercise 11.1

a.  **Prepare**:

$$
\left\{\begin{aligned}
    &H_0: \beta_{\text{schooling}} = 0 \\
    &H_1: \beta_{\text{schooling}} \neq 0
    \end{aligned}\right.
$$

Let the significance level be $\alpha = 0.05$.

**Calculate**:

$$
t^* = \frac{0.272 - 0}{0.029} \approx 9.379
$$

**Conclude**:

Since $|t^*| > Z_{0.05} = 1.96$, we reject $H_0$. There is
significant evidence that schooling affects the probability of
working for the government.

b.  The predicted probability that Friedrich will be employed by the
government is:

$$
\Phi(-4.107 + 0.272 \times 16) \approx 0.5968
$$

c.  The predicted probability that Hans will get a government job:

$$
\Phi(-4.107 + 0.272 \times 12) \approx 0.1996
$$

d.  The predicted probability that Günter will be employed by the
government:

$$
\Phi(-4.107 + 0.272 \times 24) \approx 0.9923
$$

This prediction is likely inaccurate, as the sample did not include
individuals with more than 24 years of schooling.

### Exercise 11.4

a.  In column (4), the estimated probability of being employed by the
government:

-   For men:

$$
\Phi(-1.027 - 0.242) \approx 0.1022
$$

-   For women:

$$
\Phi(-1.027) \approx 0.1522
$$

In column (5), the estimated probability:

-   For men:

$$
\frac{1}{1 + \exp[-(-1.717 - 0.455)]} \approx 0.1023
$$

-   For women:

$$
\frac{1}{1 + \exp[-(-1.717)]} \approx 0.1523
$$

In column (6), the estimated probability:

-   For men:

$$
0.152 - 0.050 = 0.102
$$

-   For women:

$$
0.152
$$

b.  Because there is only a regressor and it is binary
($\textit{Male}$), estimates for each model show the fraction of
males and females who get employed by the government. Hence the
results are close to each other.

### Exercise 11.5

a.  The predicted probability that Liam will be employed by the
government is:

$$
\Phi(-7.702 + 0.548 \times 10 + 4.352 - 0.344 \times 10) \approx 0.0951
$$

b.  The predicted probability that Anneli will be employed by the
government is:

$$
\Phi(-7.702 + 0.548 \times 12) \approx 0.1301
$$

c.  **Prepare**:

$$
\left\{\begin{aligned}
    &H_0: \beta_{\textit{male} \times \textit{schooling}} = 0 \\
    &H_1: \beta_{\textit{male} \times \textit{schooling}} \neq 0
    \end{aligned}\right.
$$

Let the significance level be $\alpha = 0.05$.

**Calculate**:

$$
t^* = \frac{-0.344 - 0}{0.096} \approx -3.583
$$

**Conclude**:

Since $|t^*| > Z_{0.05} = 1.96$, we reject $H_0$. There is
significant evidence that the effect of schooling on government
employment depends on gender.

## Computer Exercises

[(insurance.csv)](insurance.csv){target="_blank" rel="noopener"}

``` r
# remove pre-load data, and import package
rm(list = ls())
library(tidyverse)
library(estimatr) # for robust SE
library(texreg) # compile regression result into one table
library(sandwich) # Heteroskedasticity-Robust Standard Errors in testing (vcovHC)
library(car) # short for "Companion to Applied Regression", used for hypothesis testing

insurance <- read.csv("data/insurance.csv")
```

```{r}
#| echo: false
# remove pre-load data, and import package
rm(list = ls())
library(tidyverse)
library(estimatr) # for robust SE
library(texreg) # compile regression result into one table
library(sandwich) # Heteroskedasticity-Robust Standard Errors in testing (vcovHC)
library(car) # short for "Companion to Applied Regression", used for hypothesis testing
library(here)

insurance <- read.csv(here("data", "insurance.csv"))
```

### Linear Model

Consider the following specification:

$$
\textit{insured}_i = \beta_0 + \beta_1 \textit{selfemp}_i + u_i
$$

we propose the hypothesis:

$$
\left\{\begin{aligned}
&H_0: \beta_1 = 0 \\
&H_1: \beta_1 \neq 0
\end{aligned}\right.
$$

``` r
reg_1a <- lm_robust(
  formula = insured ~ selfemp,
  data = insurance,
  se_type = "HC1"
)
summary(reg_1a)
```

```{r}
#| echo: false

reg_1a <- lm_robust(
  formula = insured ~ selfemp,
  data = insurance,
  se_type = "HC1"
)
summary(reg_1a)
```

From the testing result, we have its $p$-value is very small ($<5\%$).
There is significant evidence that the self-employed are less likely to
have health insurance than wage earners.

### Non-Linear Models Comparison

Consider the following specification:

$$
\textit{insured}_i = \beta_0 + \beta_1 \textit{selfemp}_i + \sum_{k\in\mathcal{I}}\beta_k x_k + u_i
$$

where $\mathcal{I}$ denotes the set of control variables. We propose the
hypothesis:

$$
\left\{\begin{aligned}
&H_0: \beta_1 = 0 \\
&H_1: \beta_1 \neq 0
\end{aligned}\right.
$$

``` r
# complete specification
formula_1b <- insured ~ selfemp + age + deg_nd + deg_ged + deg_hs + deg_ba + deg_ma + deg_phd +
  familysz + race_bl + race_wht + male + married

# estimations
LPM_1b <- lm_robust(
  formula = formula_1b,
  data = insurance,
  se_type = "HC1"
)
Logit_1b <- glm(
  formula = formula_1b,
  data = insurance,
  family = binomial(link = "logit")
)
Probit_1b <- glm(
  formula = formula_1b,
  data = insurance,
  family = binomial(link = "probit")
)

# compile into one result table
screenreg(
  list(LPM_1b, Logit_1b, Probit_1b),
  include.ci = FALSE,
  digits = 4,
  custom.model.names = c("LPM", "Logit", "Probit")
)
```

```{r}
#| echo: false

# complete specification
formula_1b <- insured ~ selfemp + age + deg_nd + deg_ged + deg_hs + deg_ba + deg_ma + deg_phd +
  familysz + race_bl + race_wht + male + married

# estimations
LPM_1b <- lm_robust(
  formula = formula_1b,
  data = insurance,
  se_type = "HC1"
)
Logit_1b <- glm(
  formula = formula_1b,
  data = insurance,
  family = binomial(link = "logit")
)
Probit_1b <- glm(
  formula = formula_1b,
  data = insurance,
  family = binomial(link = "probit")
)

# compile into one result table
screenreg(
  list(LPM_1b, Logit_1b, Probit_1b),
  include.ci = FALSE,
  digits = 4,
  custom.model.names = c("LPM", "Logit", "Probit")
)
```

From the testing result, we have the $p$-value is very small ($<5\%$)
and reject the null hypothesis. There is significant evidence that the
self-employed are less likely to have health insurance than wage
earners.

### Prediction: LPM

``` r
person <- data.frame(
  age = 50, # 50 year-old
  race_bl = 0, race_wht = 1, # white
  male = 0, # female
  deg_nd = 0, deg_ged = 0, deg_hs = 0, deg_ba = 0, deg_ma = 0, deg_phd = 1, # phd
  familysz = 1, # family size 1
  married = 1, # married
  selfemp = 0 # wage earner
)
predict(LPM_1b, person)
```

```{r}
#| echo: false

person <- data.frame(
  age = 50, # 50 year-old
  race_bl = 0, race_wht = 1, # white
  male = 0, # female
  deg_nd = 0, deg_ged = 0, deg_hs = 0, deg_ba = 0, deg_ma = 0, deg_phd = 1, # phd
  familysz = 1, # family size 1
  married = 1, # married
  selfemp = 0 # wage earner
)
predict(LPM_1b, person)
```

From LPM, the predicted probability of having health insurance for a 50
year old white female, with a phd degree, family size of 1, married and
wage earner is $1.107$, which does not make sense because the
probability must be between 0 and 1 by definition.

### Prediction: Logit & Probit

``` r
predict(Logit_1b, person, type = "response")
```

```{r}
#| echo: false

predict(Logit_1b, person, type = "response")
```

``` r
predict(Probit_1b, person, type = "response")
```

```{r}
#| echo: false

predict(Probit_1b, person, type = "response")
```

The predicted probability of having health insurance for the same person
is 0.983 for the logit model, and 0.989 for the probit model. These
estimates make sense and differ from LPM since we use the link function
to transfer suitable function form to evaluate.

### Hypothesis Testing

We propose the following hypothesis:

$$
\left\{\begin{aligned}
&H_0: \beta_{\textit{race\_bl}} = \beta_{\textit{race\_wht}} = 0 \\
&H_1: \text{Not all coefficients are zero}
\end{aligned}\right.
$$

and perform the F-test:

``` r
linearHypothesis(
  Logit_1b,
  c("race_bl = 0", "race_wht = 0"),
  vcov = vcovHC,
  type = "HC1",
  test = "F"
)
```

```{r}
#| echo: false

linearHypothesis(
  Logit_1b,
  c("race_bl = 0", "race_wht = 0"),
  vcov = vcovHC,
  type = "HC1",
  test = "F"
)
```

From the testing result, we have the $p$-value is very small ($<5\%$)
and reject the null hypothesis. There is significant evidence that
black, white and others are not equally likely to be insured.
