---
title: 'Homework 6 Answers'
fig-cap: false
editor:
  markdown:
    wrap: 72
---

-   If you find any typos or errors, please feel free to contact me via
    email at r13323002\@ntu.edu.tw. I would appreciate it and will
    correct the solution to prevent any misunderstandings.

-   For some R commands marked with ‘OPTIONAL,’ you don’t need to learn
    them. You will be completely fine on the computer quiz even if you
    choose to ignore them.

-   If you’re having trouble understanding the code, don’t hesitate to
    ask me for help. I’ll do my best to assist you during office hours
    or after the TA session.

## Textbook Exercises

### Exercise 12.8

a.  Given the simultaneous equations,

$$
\left\{\begin{aligned}
    &Q^s_i = \beta_0 + \beta_1 P_i + u^s_i \\
    &Q^d_i = \gamma_0 + u^d_i
    \end{aligned}\right.
$$

by equilibrium condition $Q^s_i = Q^d_i \overset{\Delta}{=} Q_i$, we
get:

$$
\beta_0 + \beta_1 P_i + u^s_i = \gamma_0 + u^d_i
$$

Solving for $P_i$ and $Q_i$:

$$
\begin{aligned}
    &P_i = \frac{\gamma_0 + u^d_i - \beta_0 - u^s_i}{\beta_1} \\
    &Q_i = \gamma_0 + u^d_i
    \end{aligned}
$$

Since:

$$
\text{Cov}(P, u^s)
    = \text{Cov}\left(\frac{\gamma_0 + u^d - \beta_0 - u^s}{\beta_1}, u^s\right)
    = \text{Cov}\left(\frac{-u^s}{\beta_1}, u^s\right)
    = -\frac{1}{\beta_1}\text{Var}(u^s)
$$

We have $\text{Corr}(P, u^s) \neq 0$ and conclude that the price
variable is endogenous.

b.  We can take the probability limit on the OLS estimator:

$$
\hat{\beta}_1
    = \frac{\sum_{i=1}^{n} (P_i - \bar{P})(Q_i - \bar{Q})}{\sum_{i=1}^{n} (P_i - \bar{P})^2}
    = \frac{\frac1n \sum_{i=1}^{n} (P_i - \bar{P})(Q_i - \bar{Q})}{\frac1n \sum_{i=1}^{n} (P_i - \bar{P})^2}
    \overset{p}{\to} \frac{\text{Cov}(P, Q)}{\text{Var}(P)}
$$

and note that:

$$
\frac{\text{Cov}(P, Q)}{\text{Var}(P)}
    = \frac{\text{Cov}(P, \beta_0 + \beta_1P + u^s)}{\text{Var}(P)}
    = \beta_1 + \frac{\text{Cov}(P, u^s)}{\text{Var}(P)}
$$

Since $\text{Cov}(P, u^s) \neq 0$, we conclude that $\hat{\beta}_1$
is inconsistent.

c.  $\gamma_0$ can be estimated by OLS easily since it's equivalent to
the sample mean of $Q$.

However, to estimate $\beta_0$ and $\beta_1$, we need an
instrumental variable, which should be correlated with $P$ but
uncorrelated with $u^s$. In this case, $Q$ can serve as the
instrument because demand is completely inelastic (so that $Q$ is
not affected by shifts in supply).

### Exercise 12.9

a.  While it is plausible that having more secure property rights
affects a country's level of income, it is also plausible that
richer countries can afford institutions that protect property
rights. Thus, there is a problem of simultaneous causality bias.
According to this theory, as richer countries can afford secure
property rights, OLS estimates might be overestimated.

b.  Denote the instrumental variable settler mortality as $z$. We need
to check the following conditions:

-   **Exogeneity**: settler mortality ($z$) is not directly related
to a country's recent income ($u$). Thus, $z$ satisfies
$\text{Cov}(z, u) = 0$.
-   **Relevance**: settler mortality ($z$) affected European
settlement, and subsequently led to the establishment of early
institutions for securing property rights($x$), which have
persisted until the present day. Hence, $z$ satisfies
$\text{Cov}(z, x) \neq 0$.

### Exercise 12.10

By taking probability limit on the TSLS estimator:

$$
\hat{\beta}^{TSLS}_1
= \frac{\sum_{i=1}^{n} (Z_i - \bar{Z}) (Y_i - \bar{Y})}{\sum_{i=1}^{n} (Z_i - \bar{Z}) (X_i - \bar{X})}
= \frac{\frac1n \sum_{i=1}^{n} (Z_i - \bar{Z}) (Y_i - \bar{Y})}{\frac1n \sum_{i=1}^{n} (Z_i - \bar{Z}) (X_i - \bar{X})}
= \overset{p}{\to} \frac{\text{Cov}(Z, Y)}{\text{Cov}(Z, X)}
$$

and note that:

$$
\frac{\text{Cov}(Z, Y)}{\text{Cov}(Z, X)}
= \frac{\text{Cov}(Z, \beta_0 + \beta_1 X + \beta_2 W + u)}{\text{Cov}(Z, X)}
= \beta_1 + \frac{\beta_2 \text{Cov}(Z, W)}{\text{Cov}(Z, X)}
$$

a.  If $\text{Cov}(Z, W) \neq 0$, the IV estimator is not consistent.
The first student is correct.

b.  If $\text{Cov}(Z, W) = 0$, the IV estimator is consistent. The
second student is correct.

### Exercise 13.3

a.  The average treatment effect on test scores is

$$
\bar{X}_{\text{treatment}} - \bar{X}_{\text{control}} = 1348 - 1395 = -47
$$

b.  **Prepare**:

$$
\left\{\begin{aligned}
    &H_0: p^{\text{treatment}}_{\text{men}} = 0.5 \\
    &H_1: p^{\text{treatment}}_{\text{men}} \neq 0.5
    \end{aligned}\right.
$$

Let the significance level be 5%.

**Calculate**:

$$
t^* = \frac{\hat{p} - p_0}{\sqrt{\frac{p_0 (1 - p_0)}{n}}} = \frac{0.6 - 0.5}{\sqrt{\frac{0.5 (1 - 0.5)}{100}}} = 2
$$

**Conclude**:

Since $|t^*| > Z_{0.025} = 1.96$, we reject $H_0$. There is a
significant evidence of nonrandom assignment.

### Exercise 13.7

From the population regression, we know that:

$$
\begin{aligned}
Y_{i2} &= \alpha_i + \beta_1 X_{i2} + \beta_2 W_i + \beta_0 + v_{i2} \\
Y_{i1} &= \alpha_i + \beta_1 X_{i1} + v_{i1}
\end{aligned}
$$

By taking differences, we can obtain:

$$
\underbrace{Y_{i2} - Y_{i1}}_{\Delta Y_i}
= \beta_0 + \beta_1 \underbrace{(X_{i2} - X_{i1})}_{\Delta X_i} + \beta_2 W_i + \underbrace{(\nu_{i2} - \nu_{i1})}_{\Delta \nu_i}
$$

This eliminates the individual fixed effects ($\alpha_i$), allowing for
consistent estimation of $\beta_1$. Additionally, note that we can
define $X_i \equiv \Delta X_i$ since it's actually binary, so this can
be rewritten as:

$$
\Delta Y_i = \beta_0 + \beta_1 X_i + \beta_2 W_i + \Delta \nu_i
$$

which is equation (13.6) in the case of single $W$ regressor.

## Computer Exercises

[(card.csv)](card.csv){target="_blank" rel="noopener"}

``` r
# remove pre-load data, and import package
rm(list = ls())
library(tidyverse)
library(texreg) # compile regression result into one table
library(car) # "Companion to Applied Regression", used for hypothesis testing
library(AER) # "Applied Econometrics with R", used for IV regression

# import data
card <- read.csv("data/card.csv")
```

```{r}
#| echo: false
# remove pre-load data, and import package
rm(list = ls())
library(tidyverse)
library(texreg) # compile regression result into one table
library(car) # "Companion to Applied Regression", used for hypothesis testing
library(AER) # "Applied Econometrics with R", used for IV regression
library(here)

# import data
card <- read.csv(here("data", "card.csv"))
```

### OLS Model

``` r
reg_1a <- lm(
  formula = lwage ~ educ + south + married + exper + I(exper^2),
  data = card
)
summary(reg_1a)
```

```{r}
#| echo: false

reg_1a <- lm(
  formula = lwage ~ educ + south + married + exper + I(exper^2),
  data = card
)
summary(reg_1a)
```

### OLS Model: Add Controls

``` r
reg_1b <- lm(
  formula = lwage ~ educ + south + married + exper + I(exper^2) + iq,
  data = card
)
summary(reg_1b)
```

```{r}
#| echo: false

reg_1b <- lm(
  formula = lwage ~ educ + south + married + exper + I(exper^2) + iq,
  data = card
)
summary(reg_1b)
```

### OLS Model: Add IV

``` r
reg_1c <- lm(
  formula = educ ~ south + married + exper + I(exper^2) + nearc4,
  data = card
)
summary(reg_1c)
```

```{r}
#| echo: false

reg_1c <- lm(
  formula = educ ~ south + married + exper + I(exper^2) + nearc4,
  data = card
)
summary(reg_1c)
```

``` r
linearHypothesis(
  reg_1c,
  c("nearc4 = 0"),
  test = "F"
)
```

```{r}
#| echo: false

linearHypothesis(
  reg_1c,
  c("nearc4 = 0"),
  test = "F"
)
```

From the testing result, since F value is greater than 10,
$\textit{nearc4}$ is not a weak instrument.

### IV: Manually 2SLS

``` r
# predicted value in First stage
card$educhat <- reg_1c$fitted.values

# replace the educ by educhat
# the SEs in second stage are incorrect !!
reg_1d <- lm(
  formula = lwage ~ south + married + exper + I(exper^2) + educhat,
  data = card
)
summary(reg_1d)
```

```{r}
#| echo: false

# predicted value in First stage
card$educhat <- reg_1c$fitted.values

# replace the educ by educhat
# the SEs in second stage are incorrect !!
reg_1d <- lm(
  formula = lwage ~ south + married + exper + I(exper^2) + educhat,
  data = card
)
summary(reg_1d)
```

### IV: 2SLS by Package

``` r
# IV estimation by using package
reg_1e <- ivreg(
  formula = lwage ~ educ + south + married + exper + I(exper^2)
                         | south + married + exper + I(exper^2) + nearc4,
  data = card
)

# regression table
screenreg(
  list(reg_1a, reg_1e),
  include.ci = FALSE,
  digit = 4,
  custom.model.names = c("OLS", "IV")
)
```

```{r}
#| echo: false

# IV estimation by using package
reg_1e <- ivreg(
  formula = lwage ~ educ + south + married + exper + I(exper^2)
                         | south + married + exper + I(exper^2) + nearc4,
  data = card
)

# regression table
screenreg(
  list(reg_1a, reg_1e),
  include.ci = FALSE,
  digit = 4,
  custom.model.names = c("OLS", "IV")
)
```

### IV: Conditions

There are two conditions:

-   Relevance: $\textit{motheduc}$ and $\textit{fatheduc}$ should be
correlated with $\textit{educ}$.

-   Exogeneity: $\textit{motheduc}$ and $\textit{fatheduc}$ should be
uncorrelated with the error term.

### IV: Estimation

``` r
reg_1g <- ivreg(
  formula = lwage ~ educ + south + married + exper + I(exper^2)
                         | south + married + exper + I(exper^2) + nearc4 + fatheduc + motheduc,
  data = card
)
```

```{r}
#| echo: false

reg_1g <- ivreg(
  formula = lwage ~ educ + south + married + exper + I(exper^2)
                         | south + married + exper + I(exper^2) + nearc4 + fatheduc + motheduc,
  data = card
)
```

### IV: Overidentification Test

``` r
# residuals
resi <- reg_1g$residuals

# calculate J statistics by partial F test of residual on IVs
reg_1h <- lm(
  formula = resi ~ south + married + exper + I(exper^2) + nearc4 + fatheduc + motheduc,
  data = card
)
test_1h <- linearHypothesis(
  reg_1h,
  c("nearc4 = 0", "fatheduc = 0", "motheduc = 0"),
  test = "F"
)

# Note that J = m * F ~ chisquared(m-k), where k = #(endogeneous variables) and m = #(instrumental variables)
J <- 3 * test_1h[2, 5]
pchisq(J, df = 3-1, lower.tail = FALSE)
```

```{r}
#| echo: false

# residuals
resi <- reg_1g$residuals

# calculate J statistics by partial F test of residual on IVs
reg_1h <- lm(
  formula = resi ~ south + married + exper + I(exper^2) + nearc4 + fatheduc + motheduc,
  data = card
)
test_1h <- linearHypothesis(
  reg_1h,
  c("nearc4 = 0", "fatheduc = 0", "motheduc = 0"),
  test = "F"
)

# Note that J = m * F ~ chisquared(m-k), where k = #(endogeneous variables) and m = #(instrumental variables)
J <- 3 * test_1h[2, 5]
pchisq(J, df = 3-1, lower.tail = FALSE)
```

Because the $p$-value is greater than 0.05, under the 5% significance level,
we do not reject $H_0$. So we conclude that all instruments are exogenous.
