---
title: "Week 1 Answers (0224)"
---

## Project

1. Create an RStudio project inside a folder with proper names (e.g. TA_Session).  

2. Create folders for R scripts and data.

3. For the following questions, your code should follow the coding styles in the slides.

::: {.callout-tip title="Hint"}
It is recommended to create a new project different from last semester!
:::

## Assign Values

Create two variables called `x` and `y`, and assign values `20260224` and `409`. Compute `x / y`.
``` r
x <- 20260224
y <- 409
print(x / y)
```

```{r}
#| echo: false      

x <- 20260224
y <- 409
print(x / y)
```

## Help

Use the `help()` function to check docs, e.g. `data.frame()` or `mean()`.
``` r
help(data.frame)
help(mean)
```

## Matrix and DataFrame

1. Create a vector `name_vec` with 3 names James, Noc, and Monica.

2. Create a vector `age_vec` that stores their age 25, 30, and 28.

3. Create a 3 × 2 matrix `name_age_mat` with `name_vec` and `age_vec`, then print it in the console.

4. Create a dataframe `name_age_df` using `name_vec` and `age_vec` with column names `name` annd `age`, then print it in the console.

5. What’s the data type of `name_age_df` and `age_vec`?

``` r
name_vec <- c("James", "Noc", "Monica")
age_vec <- c(25, 30, 28)

name_age_mat <- matrix(c(name_vec, age_vec), nrow = 3)
print(name_age_mat)
class(name_age_mat)

name_age_df <- data.frame(name = name_vec, 
                          age = age_vec)
print(name_age_df)
class(name_age_df)
```


```{r}
#| echo: false      

name_vec <- c("James", "Noc", "Monica")
age_vec <- c(25, 30, 28)

name_age_mat <- matrix(c(name_vec, age_vec), nrow = 3)
print(name_age_mat)
class(name_age_mat)

name_age_df <- data.frame(name = name_vec, 
                          age = age_vec)
print(name_age_df)
class(name_age_df)
```

